---
title: "HW 3 Instructions"
author: "Penelope Pooler Eisenbies"
date: "HW 3 is Due 2/7/2024"
output: 
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
---

```{r setup, include=F}

# this line specifies options for default options for all R Chunks
knitr::opts_chunk$set(echo=T, highlight=T)

# suppress scientific notation
options(scipen=100)

# install helper package (pacman) if needed
if (!require("pacman")) install.packages("pacman", repos = "http://lib.stat.cmu.edu/R/CRAN/")

# install and load required packages
# pacman should be first package in parentheses and then list others
pacman::p_load(pacman, tidyverse, magrittr, gridExtra, lubridate, 
               knitr, kableExtra)

# verify packages 
p_loaded()

```


```{r data cleaning for HW 3, echo=F, include=F, message=FALSE}

bom22 <- read_csv("box_office_mojo_2022.csv", skip=11, show_col_types = FALSE) |>
  filter(!is.na(Day)) |>
  select(Date, `Top 10 Gross`, Gross, Releases, `#1 Release`) |>
  rename(date=Date, top10gross = `Top 10 Gross`, 
         num_releases=Releases, num1gross=Gross, num1 = `#1 Release`) |>
  
  mutate(top10gross = gsub(x=top10gross, pattern="$", replacement="", fixed=T),
         top10gross = gsub(x=top10gross, pattern=",", replacement="", fixed=T) |> as.numeric(),
         num1gross = gsub(x=num1gross, pattern="$", replacement="", fixed=T),
         num1gross = gsub(x=num1gross, pattern=",", replacement="", fixed=T) |> as.numeric()) |>
  
  mutate(date = dmy(paste(date,"2022")),
         month = month(date, label=T, abbr=T),
         day = wday(date, label=T, abbr=T),
         quart = quarter(date)) |>
  select(date, month, day, quart, top10gross:num1) |>  # select and reorder variables
  glimpse() |>                                         # examine data
  write_csv("Box_Office_Mojo_Week3_HW3.csv")           # export using write_csv


```




## Purpose

This assignment will give you experience with:

- Creating an R Project Directory and a code_data_output folder (Review)

- Creating, saving, using an R Markdown file (Review)

- Importing data

- Knitting an R Markdown file (Review)

- Creating a README file (Review)

- Using the **`dplyr`** commands along with commands to reshape and summarize data

- Creating plots with some formatting


## Instructions

### HW 3 - First Steps

**`0.`** You are encouraged to work with others and and experiment with sharing code (.Rmd files), but each student should create their own project and .Rmd file to submit.

**`1.`** Create an R project named **`HW 3 <first name> <last name>`**
   
**`2.`** Create a folder in your R project labeled **`code_data_output`**.

**`3.`** Download the provided file, **HW3_Template.Rmd** from Blackboard

**`4.`** Save the downloaded **`HW3_Template.Rmd`** file to your code_data_output folder

- Change file name to be **`HW3_FirstName_LastName.Rmd`**

- For example, I would change the template file to be named **`HW3_Penelope_Pooler.Rmd`**
   
- Change title in the file header to be 'HW 3'
   
- Specify yourself as the author
   
<br>

**`5`** Save the provided data file **`Box_Office_Mojo_Week3_HW3.csv`** into your code_data_output folder of this project.

- Note: that these data are already **clean** and useable but they will be modified in the steps below.

<br>

**NOTES** 

- Provided header text below shows the correct format.

- This header text also creates a floating Table of Contents (toc).
   
```
---
title: "HW 3"
authors: "Penelope Pooler"
date: "Due 2/7/2024"
output: 
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
---

```

**`6.`** Modify the **`setup`** chunk 

- The **`setup`** chunk text shown below has packages and setup options for HW 3.

- Copy and paste text into the setup chunk of your HW 3 Markdown (**`.Rmd`**) file that you created from the provided template.
   
<br>

```
# this line specifies options for default options for all R Chunks
knitr::opts_chunk$set(echo=T, highlight=T)

# suppress scientific notation
options(scipen=100)

# install helper package (pacman) if needed
if (!require("pacman")) install.packages("pacman", repos = "http://lib.stat.cmu.edu/R/CRAN/")

# install and load required packages
# pacman should be first package in parentheses and then list others
pacman::p_load(pacman, tidyverse, magrittr, gridExtra, lubridate, 
               knitr, kableExtra)

# verify packages 
p_loaded()

```
<br>

**`7.`** Click the green triangle or type Ctrl/Cmd + Shift + Enter to run this setup chunk.

---

### HW 3 - Part 1

#### Chunk 2: Import and Modify Categorical Variables

**`0.`** Import `Box_Office_Mojo_Week3_HW3.csv` and save it as `mojo_22` using the provided R code in Chunk 2.

**`1.`** Remove **`#`** (comment indicator) from in front of the **`select`** command and use **`!`** to omit the text variable, **`num1`**.

**`2.`** Remove **`#`** (comment indicator) from in front of incomplete code lines in the lower part of this chunk.

**`3.`** Use examples that show how to create a factor variable to complete code with **`mutate`** statement to create factor variable **`quartF`** from **`quart`**: 
- levels: `c(1,2,3,4)`
- labels: `c("1st Qtr", "2nd Qtr", "3rd Qtr", "4th Qtr")` 

<br>
   
**`4.`** Examine data with **`glimpse`** and answer Blackboard Question:

<br>

##### **BB Question 1**

After completing Part 1, the dataset, mojo_22_mod has

`____` rows and 

`____` columns

that include a date variable and

`____` character <chr> variables,

`____` numeric <dbl> variables, and

`____` factor <fct> variables.

<br>
   
---

### HW 3 - Part 2

#### Chunk 3: Modify and Create Numerical Variables

**`1.`** Remove **`#`** (comment indicator) from in front of incomplete code lines in the lower portion of this chunk. 

**`2.`** Use **`mutate`** to do the following:

- Coerce Number of Releases to be an integer variable:
     - `num_releases = as.integer(num_releases)`
     
<br>

- Create `num1_pct = num1gross/top10gross*100` and round to 2 decimal places
     - see example above
     
<br>

**`3.`** Answer Blackboard Question:


##### **BB Question 2**

The correct command used to convert a numeric variable to an integer variable is

`_____`.

When you **`glimpse`** the data after completing Part 2 (Chunk 3), the type for the num_releases variable is shown as 

`____` instead of `<dbl>`.

<br>

---

### HW 3 - Part 3

#### Chunk 4: Group and Summarize Data

**`1.`** Remove **`#`** (comment indicator) from in front of incomplete code lines in this chunk. 

  - **Note:** The first step of selecting the variables for the summary table has been completed for you.
  
<br>

**`2.`** Complete the **`group_by`** command to group the data by **`quartF`**, **`wkdayF`**

**`3.`** Complete the **`summarize`** command to create the following summary variables. Recall that **`na.rm=T`** is used remove missing values before calculating summary statistics.

- **`max_num`**, the maximum number of releases (num_releases):
  - **`max_num = max(num_releases, na.rm=T)`**
   
- **`mean_num1grossM`**, mean of the number 1 gross (num1grossM) rounded to 2 decimal places
  - **`mean_num1grossM = mean(num1grossM, na.rm=T) |> round(2)`**

<br>

**`4.`** Answer Blackboard Question:

<br>

##### **BB Question 3**

Your grouped and summarized data have

`____` rows and

`____` columns with

`____` summary numeric variables.

<br>
   
---

### HW 3 - Part 4

#### Chunk 5: Reshape Data to Create a Table

Use the summary dataset **`mojo_qtr_smry`** you created in Part 3 (Chunk 4)

**`1.`** Remove **`#`** (comment indicator) from in front of incomplete code lines in this chunk. 

**`2.`** Use **`pivot_wider`** to create a wider table with 
   
- 1 row for each quarter and 1 column for each weekday
   
- to do this specify:
  - **`id_cols=quartF, names_from=wkdayF, values_from=mean_num1grossM`**

<br>
   
**`3.`** Create a presentation version of your table with **`kable`**

**`4.`** Fill in the blanks in the Blackboard Question:

<br>

##### **BB Question 4**

The mean daily gross ($ millions) for Fridays in each quarter was:

- 1st Qtr: $____ million

- 2nd Qtr: $____ million

- 3rd Qtr: $____ million

- 4th Qtr: $____ million

---

### HW 3 - Part 5

#### Chunk 6: Reshaping and Plotting Data

Use the wide dataset **`mojo_qtr_wide`** you created in Part 4 (Chunk 5)

**`1.`** Remove **`#`** (comment indicator) from in front of incomplete code lines in this chunk. 

**`2.`** Use **`pivot_longer`** to reshape **`mojo_qtr_wide`** to be long again with these specifications:

- **`cols=M:Su, names_to="Day", values_to="mean_num1grossM"`**
  
- **`cols=M:Su`** means column labels from M to Su (Monday to Sunday will be 1 column)

- **`names_to="Day"`** means there will be a column named **`Day`** that lists all of the days, M:Su (M, T, W, Th, F, Sa, Su)

- **`values_to="mean_num1grossM"`** means the values will be in one long column named **`mean_num1grossM`**

- **NOTE:** Reshaping the data turned our 'Day' data into a character variable (again).   - I provide the code to make it a factor variable (again).

<br>
   
**`3.`** Complete the geom_bar statement to create the bar plot as follows:
   
- **`aes(x=Qtr, y=mean_num1grossM, fill=Day)`**
   
- outside of aesthetic (aes), after the comma: 
  - **`stat="identity"`**, **`position="dodge"`**

<br>

-  **`stat="identity"`** tells R to use the data values themselves instead of default (number of observations)
- **`position="dodge"`** indicates side by side bars instead of stacked

<br> 
   
4. Answer Blackboard Questions:

<br>

##### **BB Question 5**

For each option include the quotes in your answer.

**`stat=____`** tells R to create the barplot using the numeric values in the data INSTEAD of the number of observations, which is the default.

**`position=____`** indicates that the bars should be side-by-side, instead of stacked which is the default.

<br>

##### **BB Question 6**

Based on the barplot created in Part 5 (Chunk 6), which day of the week has the LOWEST mean daily gross for the top film in all 4 quarters?

<br>

---

### HW 3 - Final Steps

**`1.`** Save your HW 3 R Markdown File (.Rmd) to your code_data_output folder

- Feel free to add addtional notes to yourself above or within each chunk. 

<br>

**`2.`** Knit your R markdown File to create an HTML file (.html).

<br>

**`3.`** Answer all Blackboard questions associated with this assignment.

- You are welcome (encouraged) to work together, each student should submit their own Blackboard assignment.

<br>

**`4.`** Create a README file using the template provided.

- The dataset **Box_Office_Mojo_Week3_HW3.csv** should be 
- saved in your code_data_output folder listed in your README under **DATA FILES**
     
<br>

**`5.`** Zip your entire Project Directory into a compressed File and submit it.

- The zipped R Project should be named **`HW 3 FirstName LastName`**.
- The zipped project directory should contain:
  - The completed README file 
  - The .Rproj file
  - The code_data_output folder that contains
    - completed markdown file (.Rmd)
    - knitted .html file
    - .csv data file

<br>

---

## Grading Criteria

**(8 pts.)** Each Blackboard question for HW 3 is worth 1 or 2 points.

<br>

**(2 pts.)** Completing **HW 3 - First Steps** as specified.  

<br>
  
**(2 pts.)** Part 1: Full credit for:

- correctly excluding **`num1`** from the dataset
- correctly creating a factor variable, **`quartF`**

<br>
    
**(2 pts.)** Part 2: Full credit for:

- correctly coercing **`num_releases`** to be an integer variable
- correctly creating the variable **`num1pct`**

<br>

**(2 pts.)** Part 3: Full credit for:

- correctly grouping and summarizing the data to create to summary variables:
- **`max_num`** and **`mean_num1grossM`** 

<br>

**(2 pts.)** Part 4: Full credit for:

- correctly using **`pivot_wider`** to reshape the data to wide format
- creating a basic presentation table using **`kable`**

<br>

**(2 pts.)** Part 5: Full credit for:

- correctly using **`pivot_longer`** to reshape the data to long format
- completing the barplot code to create a barplot

<br>

**(3 pts.)** Completing the **HW 3 - Final Steps** and correctly submitting your zipped project directory:

- 1 point for creating a correct README file
- 1 point for having the .Rmd, .html, and .csv files in your code_data_output folder
- 1 point for zipping and submitting your project correctly



